### 2016-01-29 experiment2/analysis/exp2Analysis.Rmd
Update analysis script

---
title: "KeysarMouseTracking.csv"
output: html_document
---

```{r}
setwd('~/Repos/keysar_replication/experiment1/analysis/')
preMouseDExp1 = read.csv('fullSampleMouse.csv')
preErrorDExp1 = read.csv('fullSampleError.csv')
preMessageDExp1 = read.csv('fullSampleMessage.csv')

setwd('~/Repos/keysar_replication/experiment2/data/')
preMouseDExp2 = read.csv('mouse/exp2Mouse.csv')
preErrorDExp2 = read.csv('error/exp2Error.csv')
preMessageDExp2 = read.csv('message/exp2Message.csv')
library(tidyr)
library(dplyr)
library(ggplot2)
library(lme4)
library(lmerTest)
```

Filter out bad participants

```{r}
nonCriticalMistakes <- preErrorDExp1 %>%
  # only look at mistakes on noncritical (filler) items
  filter(critical != 1) %>%
  # don't want to double-count people for messing the same thing up multiple times
  filter(attemptNum == 0) %>% 
  group_by(gameid) %>%
  tally() %>%
  # implement exclusion criteria of errors on >~10% of non-critical trials
  filter(n >= 2)

bannedIDsExp1 <- nonCriticalMistakes$gameid
numParticipantsExp1 <- (length(unique(preMouseDExp1$gameid)) -
                        length(bannedIDsExp1))

nonNativeEnglishIDs <- c("0366-3", "8676-f")
nonCriticalMistakes <- preErrorDExp2 %>%
  # only look at mistakes on noncritical (filler) items
  filter(critical != 1) %>%
  # don't want to double-count people for messing the same thing up multiple times
  filter(attemptNum == 0) %>% 
  group_by(gameid) %>%
  tally() %>%
  # implement exclusion criteria of errors on >~10% of non-critical trials
  filter(n >= 2)

mistakeIDs <- as.character(nonCriticalMistakes$gameid)
bannedIDsExp2 <- unique(c(mistakeIDs, nonNativeEnglishIDs))
numParticipantsExp2 <- (length(unique(preMouseDExp2$gameid)) -
                        length(bannedIDsExp2))
errorDExp1 <- preErrorDExp1 %>%
  filter(!(gameid %in% bannedIDsExp1))
mouseDExp1 <- preMouseDExp1 %>%
  filter(!(gameid %in% bannedIDsExp1))
messageDExp1 <- preMessageDExp1 %>%
  filter(!(gameid %in% bannedIDsExp1))

errorDExp2 <- preErrorDExp2 %>%
  filter(!(gameid %in% bannedIDsExp2))
mouseDExp2 <- preMouseDExp2 %>%
  filter(!(gameid %in% bannedIDsExp2))
messageDExp2 <- preMessageDExp2 %>%
  filter(!(gameid %in% bannedIDsExp2))
```

Make error table

```{r}
eDExp2 <- errorDExp2 %>%
  filter(attemptNum == 0) %>%
  group_by(gameid, condition) %>%
  summarize(n = sum(criticalError)) %>%
  mutate(cond1 = as.numeric(n >= 1)) %>%
  mutate(cond2 = as.numeric(n >= 2)) #

errorTableExp2 <- eDExp2 %>%
  group_by(condition) %>%
  summarise(atLeastOnce = sum(cond1)/numParticipantsExp2, 
            atLeastTwice = sum(cond2)/numParticipantsExp2,
            total = sum(n) / (4 * numParticipantsExp2))
errorTableExp2

# To illustrate how much a few items are driving the effect, we exclude three items 
# where more than 60% of participants made errors.
excludedEDExp2 <- errorDExp2 %>%
  filter(attemptNum == 0) %>%
  filter(!(objectSet %in% c(5, 6))) %>% # Just 5 & 6?
  group_by(gameid, condition) %>%
  summarize(n = sum(criticalError)) %>%
  mutate(cond1 = as.numeric(n >= 1)) %>%
  mutate(cond2 = as.numeric(n >= 2)) %>%
  group_by(condition) %>%
  summarise(atLeastOnce = sum(cond1)/numParticipantsExp2, 
            atLeastTwice = sum(cond2)/numParticipantsExp2,
            total = sum(n) / (4 * numParticipantsExp2))
excludedEDExp2
```

When we look at errors per item, we have to remember that not all items had the same number of participants in the experimental condition.

```{r}
numPlayersPerConditionExp2 = mouseDExp2 %>% 
  filter(critical == 1) %>% 
  filter(attemptNum == 0) %>% 
  group_by(gameid, condition, objectSet) %>% 
  summarize() %>% 
  group_by(condition, objectSet) %>% 
  summarize(total = n()) %>% 
  filter(condition == "exp")
```

Follow-up: are all critical items equal?

```{r}
itemInequalitiesExp2 <- errorDExp2 %>%
  filter(condition == "exp") %>%
  filter(criticalError == 1) %>%
  filter(attemptNum == 0) %>%
  group_by(objectSet) %>%
  tally() %>%
  right_join(numPlayersPerConditionExp2, by = c('objectSet')) %>%
  mutate(objectSet = objectSet,
         errorCount = n, 
         correctCount = total - n) %>%
  mutate(errorRate = errorCount / (errorCount + correctCount)) %>%
  select(objectSet, errorCount, correctCount, errorRate, total)

itemInequalitiesExp2[1,2:4] = c(0, itemInequalitiesExp2[1,]$total, 0)

itemWiseTestExp2 <- chisq.test(itemInequalitiesExp2[-c(1,4,5)])
print(itemWiseTestExp2)
```

Follow-up: Did error rates go down significantly in Exp. 2?

```{r}
exp1NumErrors = eDExp1 %>% 
  filter(condition == "exp") %>% 
  group_by(condition) %>%
  summarize(numErrors = sum(n))
exp2NumErrors = eDExp2 %>% 
  filter(condition == "exp") %>% 
  group_by(condition) %>%
  summarize(numErrors = sum(n))

exp1TotalPossible = 4 * numParticipantsExp1
exp2TotalPossible = 4 * numParticipantsExp2

numErrors = c(exp1NumErrors$numErrors, exp2NumErrors$numErrors)
totalPossible = c(exp1TotalPossible, exp2TotalPossible)
prop.test(numErrors, totalPossible)
```

Set up message data
-------------------

Need to join the coder ratings...

```{r}
library(boot)
library(irr)
turkerRatings <- read.csv("../../normingExperiment/data/keysarNorming-trials.csv")
relevantProperties <- c("objectSet", "label")

criticalMessagesExp1 <- read.csv("../analysis/messagesExp1.csv") %>%
  left_join(turkerRatings, by = relevantProperties) %>%
  mutate(exp = "Exp.1")
criticalMessagesExp2 <- read.csv("../analysis/messagesExp2.csv") %>%
  left_join(turkerRatings, by = relevantProperties) %>%
  mutate(exp = "Exp.2")

rawOverinformativeCodes <- rbind(criticalMessagesExp1, criticalMessagesExp2) %>%
  filter(!(gameid %in% bannedIDsExp2)) %>%
  filter(!(gameid %in% bannedIDsExp1))

overinformativeCodes = rawOverinformativeCodes %>%
  group_by(exp, label, objectSet, workerid, referent) %>%
  summarize(response = mean(response))

View(overinformativeCodes)
summary(lmer(response ~ exp * referent + (1 | workerid) + (1 | objectSet), 
             data = overinformativeCodes))

stat = function(data, indices) {
  d <- data[indices,]
  applicability = d %>%   
    group_by(exp, referent) %>%
    summarize(m = mean(response))
  return(applicability$m)
}

res = boot(data = overinformativeCodes, statistic = stat, R = 1000)
Exp1DistrCI = boot.ci(res, type = "basic", index=1)
Exp1targetCI = boot.ci(res, type = "basic", index=2)
Exp2DistrCI = boot.ci(res, type = "basic", index=3)
Exp2targetCI = boot.ci(res, type = "basic", index=4)

plottingDF = data.frame(referent = c("distractor", "target", "distractor", "target"),
                        exp = c("Exp.1", "Exp.1", "Exp.2", "Exp.2"),
                        response = res$t0,
                        lowerCI = c(Exp1DistrCI$basic[1,4], Exp1targetCI$basic[1,4],
                                    Exp2DistrCI$basic[1,4], Exp2targetCI$basic[1,4]),
                        upperCI = c(Exp1DistrCI$basic[1,5], Exp1targetCI$basic[1,5],
                                    Exp2DistrCI$basic[1,5], Exp2targetCI$basic[1,5]))
dodge <- position_dodge(width=0.9)
ggplot(plottingDF, aes(x = exp, y = response, fill = referent)) +
  geom_bar(stat = "identity", position = dodge) +
  geom_errorbar(aes(ymin = lowerCI, ymax = upperCI), position = dodge, width = 0.25) +
  theme_bw() +
  ylab("mean fitness of label")
ggsave("../../writing/cogsci-submission/images/fitnessInteraction.jpeg")

# spread out data, so each row is a label, and each column is a turker
wideRatings = turkerRatings %>% 
  group_by(workerid, label, referent) %>% 
  summarize(m = mean(response)) %>% 
  spread(workerid, m)

# Interrater reliability
icc(wideRatings[,-c(1,2)],
    model = "twoway", type = "agreement")
```

Compare item-wise error rates and fitness ratios

```{r}
library(binom)
Exp1Errs = itemInequalitiesExp1 %>%
  mutate(response = errorRate) %>%
  mutate(condition = "error") %>%
  do(mutate(., lower = binom.confint(.$errorCount, .$total, 
                                     method = "bayes")$lower,
               upper = binom.confint(.$errorCount, .$total, 
                                     method = "bayes")$upper)) %>%
  select(objectSet, response, condition, lower, upper)

stat = function(data, indices) {
  d <- data[indices,]
  applicability = d %>%   
    group_by(objectSet) %>%
    summarize(m = .5 * mean(distractor)/mean(target))
  return(applicability$m)
}

Exp1Fitness = overinformativeCodes %>%
  filter(exp == "Exp.1") %>%
  spread(referent, response)

res = boot(data = Exp1Fitness, statistic = stat, R = 1000)
Exp1PlottingDF = data.frame(response = rep(0, 8),
                        upper = rep(0,8),
                        lower = rep(0,8),
                        objectSet = 1:8,
                        condition = rep("fitness", 8))
for (i in 1:8) {
  ci = boot.ci(res, type = "basic", index=i)
  Exp1PlottingDF[i,]$response = res$t0[i]
  Exp1PlottingDF[i,]$lower = ci$basic[4]
  Exp1PlottingDF[i,]$upper = ci$basic[5]
}

Exp1ItemwiseForPlot = rbind(Exp1PlottingDF, Exp1Errs)
g <- (ggplot(Exp1ItemwiseForPlot, aes(x = objectSet, y = response, 
                                      group = condition, fill = condition))
  + geom_bar(stat = "identity", position = "dodge")
  + geom_errorbar(aes(ymax = upper, ymin = lower), 
                  position = dodge, width = 0.25)
  + ylab("ratio (scaled)"))
g
ggsave("../../writing/cogsci-submission/images/itemWiseFitness.jpeg")
```

Same thing for Exp 2

```{r}
library(binom)
Exp2Errs = itemInequalitiesExp2 %>%
  mutate(response = errorRate) %>%
  mutate(condition = "error") %>%
  do(mutate(., lower = binom.confint(.$errorCount, .$total, 
                                     method = "bayes")$lower,
               upper = binom.confint(.$errorCount, .$total, 
                                     method = "bayes")$upper)) %>%
  select(objectSet, response, condition, lower, upper)

stat = function(data, indices) {
  d <- data[indices,]
  applicability = d %>%   
    group_by(objectSet) %>%
    summarize(m = .25 * mean(distractor)/mean(target))
  return(applicability$m)
}

Exp2Fitness = overinformativeCodes %>%
  filter(exp == "Exp.2") %>%
  spread(referent, response)

res = boot(data = Exp2Fitness, statistic = stat, R = 1000)
Exp2PlottingDF = data.frame(response = rep(0, 8),
                        upper = rep(0,8),
                        lower = rep(0,8),
                        objectSet = 1:8,
                        condition = rep("fitness", 8))
for (i in 1:8) {
  ci = boot.ci(res, type = "basic", index=i)
  Exp2PlottingDF[i,]$response = res$t0[i]
  Exp2PlottingDF[i,]$lower = ci$basic[4]
  Exp2PlottingDF[i,]$upper = ci$basic[5]
}

Exp2ItemwiseForPlot = rbind(Exp2PlottingDF, Exp2Errs)
g <- (ggplot(Exp2ItemwiseForPlot, aes(x = objectSet, y = response, 
                                      group = condition, fill = condition))
  + geom_bar(stat = "identity", position = "dodge")
  + geom_errorbar(aes(ymax = upper, ymin = lower), 
                  position = dodge , width = 0.25)
  + ylab("measure (z-scored)"))
g
```

```{r}
# within exp. correlations high
cor(Exp1Errs$response, Exp1PlottingDF$response)
cor(Exp2Errs$response, Exp2PlottingDF$response)

# Cross exp. correlations low
cor(Exp2Errs$response, Exp1PlottingDF$response)
cor(Exp1Errs$response, Exp2PlottingDF$response)
```

Want to show whether precision helps avoid errors (mixed model?)

First, we join the error data with the critical item coding data.

```{r}
criticalErrorsExp1 = errorDExp1 %>%
  filter(attemptNum == 0) %>%
  filter(criticalError == 1) %>%
  mutate(exp = "Exp.1") %>%
  select(gameid, objectSet, exp, criticalError)
criticalErrorsExp2 = errorDExp2 %>% 
  filter(attemptNum == 0)%>% 
  filter(criticalError == 1) %>%
  mutate(exp = "Exp.2") %>%
  select(gameid, objectSet, exp, criticalError)

criticalErrors = rbind(criticalErrorsExp1, criticalErrorsExp2)

itemWiseCodes <- rawOverinformativeCodes %>% 
  left_join(criticalErrors, by = c("gameid", "objectSet", "exp")) %>%
  mutate(criticalError = ifelse(is.na(criticalError),0,1)) %>%
  group_by(gameid, objectSet, referent, criticalError, exp) %>%
  summarize(meanFitness = mean(response)) %>%
  spread(referent, meanFitness)
library(lme4)
View(itemWiseCodes)
mmRes1 <- glmer(criticalError ~ scale(target) + scale(distractor) + (1 | gameid), data = itemWiseCodes, family = binomial)
summary(mmRes1)
mmRes2 <- glmer(criticalError ~ scale(target) + scale(distractor) + factor(objectSet) + (1 | gameid),
      data = itemWiseCodes, family = binomial)

anova(mmRes1, mmRes2)

newItemWiseCodes = itemWiseCodes %>% 
  mutate(modelPreds = exp(predict(mmRes1))/(exp(predict(mmRes1))+1)) %>%
  group_by(exp, objectSet) %>%
  summarize(empError = mean(criticalError),
            modelError = mean(modelPreds))

(ggplot(newItemWiseCodes, aes(x = empError, y = modelError, color = exp)) 
   + geom_point()
   + geom_smooth(method = lm))

# Shoudln't worry too much about convergence errors: appear to be false positives
# https://groups.google.com/forum/#!topic/predicts_project/gVfnVxCSYog

gg <- mmRes2@optinfo$derivs$grad
hh <- mmRes2@optinfo$derivs$Hessian
vv <- sqrt(diag(solve(hh/2)))
print(summary(abs(gg*vv)))

relgrad <- with(mmRes2@optinfo$derivs,solve(Hessian,gradient))
print(max(abs(relgrad)))

```

Set up mouse data

```{r}
mouDataExp2 <- mouseDExp2 %>%
  filter(attemptNum == 0)

mDExp2 <- messageDExp2 %>% 
  filter(attemptNum == 0) %>%
  filter(sender == "director") %>%
  group_by(gameid, objectSet, instructionNum) %>%
  mutate(messageStamp = first(time)) %>%
  select(gameid, condition, attemptNum, instructionNum, critical, 
         objectSet,sender,contents,messageStamp) %>%
  distinct(gameid,contents)

joinedExp2 <- mouDataExp2 %>%
  right_join(mDExp2, by = c("gameid", "condition", "objectSet", "instructionNum",
                            "attemptNum", "critical")) %>%
  left_join(errorDExp2, by = c("gameid", "condition", "objectSet", "attemptNum",
                               "instructionNum", "critical"))
View(joinedExp2)
```

```{r}
dExp2 <- joinedExp2 %>% 
  filter(critical == 1) %>% 
  group_by(objectSet, condition) %>%
  filter(time > messageStamp) %>%
  mutate(mouseY = 600 - mouseY) %>%
  mutate(targetY = 600 - targetY) %>%
  mutate(begTargetX = first(targetX),
         begTargetY = first(targetY),
         distractorX = as.numeric(levels(distractorX))[distractorX],
         distractorY = 600 - as.numeric(levels(distractorY))[distractorY]) %>%
   mutate(begDistrX = first(distractorX),
          begDistrY = first(distractorY)) %>%
   filter(targetX == begTargetX) %>%
   filter(targetY == begTargetY) %>%
   filter(distractorX == begDistrX) %>%
   filter(distractorY == begDistrY) 
View(dExp2)
```

Compute hover time statistics

```{r}

sem <- function(x) {sd(x, na.rm = T) / sqrt(length(x))}
ci95 <- function(x) {sem(x) * 1.96}

heatMapForLMExp2 <- dExp2 %>%
  group_by(gameid, objectSet) %>%
  mutate(inTargetSquare = as.numeric(mouseX > targetX - 75 
                                     & mouseX < targetX + 75
                                     & mouseY > targetY - 75
                                    & mouseY < targetY + 75)) %>%
  mutate(inDistractorSquare = as.numeric(mouseX > distractorX - 75
                                         & mouseX < distractorX + 75
                                         & mouseY > distractorY - 75
                                         & mouseY < distractorY + 75)) %>%
  group_by(gameid, condition) %>%
  summarise(distractor = sum(inDistractorSquare) / n(),
         target = sum(inTargetSquare) / n()) %>%
  gather(cellType, percent, distractor, target)

heatMapForLMExp2

t1 <- lm(percent ~ condition * cellType, data = heatMapForLMExp2)
summary(t1)
```

Plot it?

```{r}
heatMapForPlotExp2 <- dExp2 %>%
  group_by(gameid, objectSet) %>%
  mutate(inTargetSquare = as.numeric(mouseX > targetX - 75 
                                     & mouseX < targetX + 75
                                     & mouseY > targetY - 75
                                    & mouseY < targetY + 75)) %>%
  mutate(inDistractorSquare = as.numeric(mouseX > distractorX - 75
                                         & mouseX < distractorX + 75
                                         & mouseY > distractorY - 75
                                         & mouseY < distractorY + 75)) %>%
  group_by(gameid, condition) %>%
  summarize(distractor = sum(inDistractorSquare) / n(),
            target = sum(inTargetSquare) / n()) %>%
  gather(cellType, percent, distractor, target) %>%
  group_by(condition, cellType) %>%
  summarize(error = sem(percent),
            percent = mean(percent))

heatMapForPlotExp2

g <- (ggplot(heatMapForPlotExp2, aes(x = condition, y = percent, 
                                     group = cellType, color = cellType)) 
      + geom_line(aes(linetype = cellType), size = 2)
      + geom_errorbar(aes(ymax = percent + error, 
                          ymin = percent - error), size =2 ,
                      width = 0.1)
     + ylim(0, .45)
      + ggtitle("Experiment 2 Hover-Time"))
g
ggsave("../../writing/cogsci-submission/images/exp2MouseTracking.pdf")
```

Compare mouse data from Exp1 w/ mouse data from Exp2:

```{r}
combinedHeatMap = rbind(heatMapForLMExp1 %>%
                          mutate(exp = "Exp.1") %>% 
                          filter(cellType == "target"),
                        heatMapForLMExp2 %>% 
                          mutate(exp = "Exp.2") %>%
                          filter(cellType == "target"))
combinedHeatMapForPlot <- combinedHeatMap %>%
  group_by(condition, exp) %>%
  summarize(error = ci95(percent),
            percent = mean(percent))

g <- (ggplot(combinedHeatMapForPlot, aes(x = exp, y = percent, 
                                  group = condition, fill = condition)) 
      + geom_bar(position = dodge, stat= "identity")
      + geom_errorbar(aes(ymax = percent + error, 
                          ymin = percent - error), position = dodge,
                          width = 0.25)
      + ylim(0, .5)
      + ggtitle("Decreased hover-time on target"))
g
combinedRes = lm(percent ~ exp * condition, data = combinedHeatMap)
summary(combinedRes)
```

```{r}
# heatMapForPlot$new_labels = as.factor(sapply(X = heatMapForPlot$objectSet, 
#                               FUN = function(v) {return(paste("Item", v))}))
# 
# g <- (ggplot(heatMapForPlot, aes(x = mouseX, y = mouseY, color = condition)) +
#       geom_vline(xintercept = c(0, 150,300,450, 600)) +
#       geom_hline(yintercept = c(0, 150,300,450, 600)) +
#       geom_point() +
#       geom_point(aes(x = begTargetX, y = begTargetY), size = 20,
#                  shape = 4, color = "black", show_guide = FALSE) +
#       geom_point(aes(x = begDistrX, y = begDistrY), size = 20,
#                  shape = 4, color = "grey50", show_guide = FALSE) + 
#       theme(panel.grid.major = element_blank(), 
#             panel.grid.minor = element_blank()) +
#       xlim(0, 600) +
#       ylim(0, 600) +
#       theme(aspect.ratio = 1) +
#       facet_wrap(~ new_labels, nrow = 2))
# g
```