### 2017-11-13 modeling/modelAnalysis.Rmd
Fix numeric issue when there are -Infs

---
title: "R Notebook"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.

```{r}
library(tidyverse)
```

```{r}
```

```{r}
sumlogprob <- function(a, b) {
  if(a>b) {
    return(a+log1p(exp(b-a)))
  } else{
    return(b+log1p(exp(a-b)))
  }
}

params <- read_csv('./bdaOutput/testingParams.csv', col_types = c('cnnnnnnd'))#, delim = ',')

View(params)

modelOutput = params %>%
  #filter(logLikelihood != -Inf) %>%
  group_by(model) %>%
  summarize(logLikelihood = reduce(logLikelihood, sumlogprob) - log(length(logLikelihood)))

print(modelOutput)
```

```{r}
params %>%
  group_by(model) %>%
  #arrange(logLikelihood) %>%
  filter(logLikelihood == max(logLikelihood))

```