### 2018-05-05 analysis/final_analysis.Rmd
Exclude buggy trials; look at ref exps

---
title: "R Notebook"
output: html_notebook
---

# General imports

```{r}
library(jsonlite)
library(tidyverse)
library(ggthemes)
library(lme4)
library(lmerTest)
library(tidyboot)
```

# Experiment 1

```{r}
raw <- read_csv('../data/final/experiment1/chatMessage/messagesFromMongo.csv')
exp1_messages <- read_csv('../data/final/experiment1/chatMessage/messagesFromMongo.csv') %>%
  mutate(critical = ifelse(critical == 'True', 1, 0)) %>%
  rename(messageTime = serverTime) %>%
  filter(sender == 'director') %>%
  group_by(gameid, trialNum, instructionNum, attemptNum, critical, condition, objectSet, targetObject) %>%
  summarize(text = paste0(contents, collapse = ' '),
            typingRT = sum(typingRT)) %>%   # Concatenate multiple utterance on same trial
  mutate(numRawWords = str_count(text, "\\S+"))

## Have to post-hoc attemptNum and instructionNum bugs (got reset for next trial prematurely)
## Also, remove trials where multi-drop bug happened (i.e. where clicking multiple times triggered multiple instruction resetting)
exp1_drops <- read_csv('../data/final/experiment1/drop/dropsFromMongo.csv') %>%
  mutate(critical = ifelse(critical == 'True', 1, 0),
         correct = ifelse(correct == 'correct', 1, 0),
         instructionNum = ifelse(correct == 0, instructionNum + 1, instructionNum)) %>%
  group_by(gameid, targetObject) %>%
  filter(length(unique(instructionNum)) < 2) %>%
  group_by(gameid, instructionNum, trialNum) %>%
  mutate(attemptNum = ifelse(attemptNum == 0, max(attemptNum), attemptNum - 1)) %>%
  rename(dropTime = serverTime) %>%
  select(-X1, -eventType)
```

```{r}
exp1_subjInfo <- read_csv('../data/final/experiment1/turk/subjInfo.csv')
incompleteIDs <- unique((exp1_messages %>% 
                           group_by(gameid, trialNum, instructionNum) %>% 
                           summarize(id = row_number()) %>% 
                           group_by(gameid) %>% 
                           tally() %>% 
                           filter(n != 32))$gameid)

confused <- unique((exp1_subjInfo %>% filter(understandsInstructions != 'yes'))$gameid)
nonNative <- unique((exp1_subjInfo %>% filter(nativeEnglish != 'yes'))$gameid)

nonCriticalMistakes <- unique((exp1_drops %>%
  # only look at mistakes on noncritical (filler) items
  filter(critical != 1) %>%
  # don't want to double-count people for messing the same thing up multiple times
  filter(attemptNum == 1) %>% 
  group_by(gameid, correct) %>%
  tally() %>%
  # implement exclusion criteria of errors on >~10% of non-critical trials
  filter(n > 1))$gameid)

buggyTrials <- read_csv('../data/final/experiment1/drop/dropsFromMongo.csv') %>%
  mutate(critical = ifelse(critical == 'True', 1, 0),
       correct = ifelse(correct == 'correct', 1, 0),
       instructionNum = ifelse(correct == 0, instructionNum + 1, instructionNum)) %>%
  group_by(gameid, targetObject) %>%
  mutate(numInstructions = length(unique(instructionNum))) %>%
  filter(numInstructions > 1)

taboo <- c('4241-cda4a373-ffd7-41e3-ae60-82b9d377a455')
onlyLocation <- c('3222-f68d2df5-8977-4ff2-9ee9-c60f8ddb61f1', '9457-4189cc9b-a82b-4db9-828d-bf90d1987aee')
exp1_badGames <- c(incompleteIDs, nonNative, confused, nonCriticalMistakes, taboo, onlyLocation, unique(buggyTrials$gameid))

exp1_d <- exp1_drops %>% 
  left_join(exp1_messages) %>% #, by = c('gameid', 'trialNum', 'instructionNum', 'trialType', 
                      #         'targetObject', 'critical', 'condition', 'attemptNum')) %>%
  filter(!(gameid %in% exp1_badGames))
```

## Descriptive stats

How many recruited, how many excluded, for what reasons, etc?

```{r}
paste0('recruited ', length(exp1_subjInfo$gameid))
paste0('total games ', length(unique(exp1_drops$gameid)))
paste0('incomplete ', length(incompleteIDs))
paste0('too many mistakes ', length(setdiff(union(confused, nonCriticalMistakes), incompleteIDs)))
paste0('non-native english ', length(setdiff(nonNative, union(confused, union(nonCriticalMistakes, incompleteIDs)))))
paste0('final sample ', length(unique(exp1_d$gameid)))
```

Happened to get slightly more people in scripted condition

```{r}
exp1_d %>% 
  group_by(gameid, condition) %>% 
  tally() %>% 
  group_by(condition) %>% 
  summarize(numInCondition = length(n))
```

And slightly uneven numbers of 'experimental' trials in each item...

```{r}
exp1_d %>% 
  group_by(gameid, objectSet, trialType, condition) %>% 
  tally() %>% 
  group_by(condition, objectSet, trialType) %>% 
  summarize(numInCondition = length(n))
```

## Listener Analyses

First, look at error rates

```{r}
critTrials <- exp1_d %>% 
  filter(attemptNum == 0) %>%
  filter(critical == 1) %>%
  filter(trialType == 'exp')

critTrials %>%
  group_by(gameid, condition) %>%
  summarize(pctCriticalError = 4 - sum(correct)) %>%
  ggplot(aes(x = pctCriticalError, fill = condition)) +
    geom_histogram(alpha = 0.5, position = 'dodge', binwidth = 0.5)  +
    facet_grid(condition ~ .) +
    theme_few() +
    ylab('# pairs making k critical errors') +
    xlab('# critical errors') + 
    theme(aspect.ratio = 1, legend.position="none") +
    scale_fill_colorblind()+
    ggtitle('critical errors')

ggsave('../writing/journal_manuscript/figures/criticalErrors.pdf')
```

```{r}
critTrials %>%
  group_by(condition) %>%
  mutate(error = 1 - correct) %>%
  tidyboot_mean(column = error) %>%
  ggplot(aes(x = condition, y =empirical_stat, fill = condition)) +
    geom_bar(stat = 'identity', alpha = .5) +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0) +
    ylab('% errors') +
    scale_fill_colorblind()+
    theme_few() +
    theme(aspect.ratio = 1, legend.position="none")

ggsave('../writing/journal_manuscript/figures/errorMeans.pdf', height = 3, width = 3)
```

```{r}
errorTable <- critTrials %>%
  group_by(gameid, condition) %>%
  summarize(numErrors = 4 - sum(correct),
            numPossible = length(correct),
         atLeastOnce = sum(correct) < 4,
         atLeastTwice = sum(correct) < 3,
          atLeastThree = sum(correct) < 2,
         pctErrors = 1 - mean(correct)) %>%
  group_by(condition) %>%
  summarize(pctAtLeastOnce = mean(atLeastOnce),
            pctAtLeastTwice = mean(atLeastTwice),
            pctAtLeastThree = mean(atLeastThree),
            avgPctErrors = mean(pctErrors),
            totalNumErrors = sum(numErrors),
            totalPossible = sum(numPossible))
errorTable
```

### Statistical tests of error reduction

```{r}
prop.test(errorTable$totalNumErrors, errorTable$totalPossible)
```

### Control for item-level error rates...

```{r}
summary(glmer(correct ~ condition + (1|gameid) + (1+condition | objectSet), family = 'binomial', 
              data = critTrials))
```

## Error heterogeneity across items...

```{r}
diffs <- critTrials %>% 
  mutate(error = 1 - correct) %>%
  group_by(objectSet, condition, targetObject) %>%
  summarize(pctErrors = mean(error)) %>%
  spread(condition, pctErrors) %>%
  mutate(errorReduction = scripted - unscripted) %>%
  gather(condition, pctErrors, scripted, unscripted)

critTrials %>% 
  left_join(diffs) %>%
  mutate(error = 1 - correct) %>%
  group_by(objectSet, condition, targetObject, errorReduction) %>%
  tidyboot_mean(column = error) %>%
  ggplot(aes(x = reorder(targetObject, -errorReduction), y = empirical_stat, color = condition, group = condition)) +
    geom_point() +
    geom_line(stat = 'identity', position = 'dodge') +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0) +
    theme_few() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ggtitle('error heterogeneity across item') +
    xlab('critical item') +
    ylab('% error in condition')

ggsave('../writing/journal_manuscript/figures//itemHeterogeneity.pdf')
```

```{r}
exp1_d %>% 
  filter(attemptNum == 0) %>%
  filter(correct == 1) %>%
  filter(critical == 1) %>% 
  ungroup() %>%
  filter(responseTime < median(responseTime, na.rm = T) +  3* sd(responseTime)) %>%
  group_by(trialType, condition) %>%
  mutate(m = median(responseTime)) %>%
  ggplot(aes(x = responseTime)) +
    geom_histogram() +
    geom_vline(aes(xintercept = m))+
    facet_wrap(trialType ~ condition) +
    scale_x_log10() +
    theme_few() 
```

```{r}
summary(lmer(log(responseTime) ~ trialType * condition + (1 | gameid),
     data = exp1_d %>% 
        filter(attemptNum == 0) %>%
        filter(correct == 1) %>%
        filter(critical == 1) %>% 
        ungroup() %>%
        filter(responseTime < median(responseTime, na.rm = T) +  3* sd(responseTime))))
```

## Mouse-tracking

Even on critical trials where they didn't make errors, how often did they 'consider' the distractor (i.e. hover over the distractor cell)

```{r}
d.onDistractor <- exp1_d %>% 
  left_join(read_csv('../data/final/experiment1/mouseOverDistractor/distractorMouseFromMongo.csv') %>% 
              mutate(critical = ifelse(critical == 'True', 1, 0))
            ) %>%
  select(-serverTime, -X1) %>%
  filter(attemptNum == 0) %>%
  filter(correct == 1) %>%
  filter(critical == 1) %>%
  group_by(gameid, trialNum) %>%
  mutate(toggle = floor((row_number() - 1) / 2)) %>%
  mutate(onOff = ifelse(is.na(onOff), 'off', onOff)) %>%
  spread(onOff, timeElapsed) %>%
  mutate(diff = ifelse(is.na(off), 0, off - on)) %>%
  group_by(gameid,condition,objectSet, trialType) %>%
  summarize(totalAmtTimeOnDistractor = sum(diff)) %>%
  mutate(hovered = totalAmtTimeOnDistractor > 0) %>%
  ungroup()

dodge <- position_dodge(width=0.9)
d.onDistractor %>%
  group_by(condition, trialType) %>%
  tidyboot_mean(column = hovered) %>%
  ggplot(aes(x = condition, y=empirical_stat, fill = trialType)) +
    geom_bar(stat = 'identity', position = dodge, alpha =.5) +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0, position = dodge) +
    theme_few() +
    ylab('% trials hovering over distractor') +
    scale_fill_colorblind()+
    theme(aspect.ratio = 0.5)

ggsave('../writing/journal_manuscript/figures//hoverOverDistractor.pdf', height = 3, width = 6)
```

```{r}
summary(glmer(hovered ~ trialType * condition + (1 + trialType|gameid) + (1 |objectSet), data = d.onDistractor, family = 'binomial'))
```

Note: because I stopped recording from the mouse as soon as someone starts dragging, it's technically possible for them to hover over the distractor object, click and start to drag it, but then drop it in place and go over to the target object. So they get it 'correct' even though it looks like their mouse never gets close...

Note: a couple researcher degrees of freedom here: 
* do we take the time from 'object reveal' or from first mouse move (i.e. subtract off the first time recorded NOT in target or distractor)? 
* do we restrict to only 'correct' trials (i.e. to make an argument like "even when they got it right, they were still distracted") or include everything?
* how big a radius around the target center should we consider a 'mouse-fixation' (obviously needs to be close enough to move it, but objects were different sizes...)

```{r}
# Touch people?
touch <- c('0095-03bb67a7-ae89-402d-9f7f-c5898784e63f', '0093-b4e1ce5c-2e5c-462a-bfed-87fa81611139')
targetRadius = 175
d.mouse <- exp1_d %>% ungroup() %>%
  left_join(read_csv('../data/final/experiment1/updateMouse/mouseFromMongo.csv') %>%
                  mutate(critical = ifelse(critical == 'True', 1, 0))
  ) %>%
  filter(!(gameid %in% touch)) %>%
  filter(critical == 1) %>% filter(attemptNum == 0) %>% filter(correct == 1) %>%
  select(-critical, -attemptNum, -eventType, -iterationName) %>%
  mutate(distractorDistance = as.numeric(distractorDistance)) %>%
  group_by(condition, trialType, gameid, objectSet, onTarget = targetDistance < targetRadius) %>%
  summarize(firstTime = first(timeFromReveal), 
            lastTime = last(timeFromReveal)) %>%
  group_by(condition, trialType, gameid, objectSet) %>%
  summarize(initMouseMove = min(firstTime), 
            initTargetHover = max(firstTime) - initMouseMove, 
            finalTargetHover = max(lastTime) - initMouseMove) %>%
  select(-initMouseMove) %>%
  gather(metric, value, initTargetHover, finalTargetHover) %>%
  #spread(onTarget, value) %>%
  #mutate(timeFromFirstMouseMove = `TRUE` - `FALSE`) %>%
  #ungroup() %>%
  #filter(firstTime < mean(firstTime, na.rm = T) + 3 * sd(firstTime, na.rm = T)) %>%
  group_by(condition, trialType, metric) %>%
  tidyboot_mean(column = value)

d.mouse
ggplot(d.mouse, aes(x = trialType, y = empirical_stat, fill = metric)) +
  geom_bar(stat= 'identity', position = dodge) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), position = dodge, width = 0) +
  facet_wrap(~ condition) +
  theme_few()
```

## Prep informativity norming stims

```{r}
refExps <- exp1_d %>% 
  filter(condition == 'unscripted') %>% 
  filter(critical == 1) %>% 
  filter(attemptNum == 0) %>% 
  ungroup() %>%
  select(gameid, objectSet, text)

write_csv(refExps, './refExpressions_raw.csv')
```

For 6 of the 8 items, unscripted speakers had significantly more words in their referring expressions... (n.s. for the other 2, but idiosyncratic)

```{r}
refs <- read_csv('../data/final/experiment1/refExpressions.csv')
scripted <- data.frame(objectSet = c(1,2,3,4,5,6,7,8), 
                       m = c('glasses', 'bottom block', 'tape', 'large measuring cup', 'brush', 'eraser', 'small candle', 'mouse'),
                       numRawWords = c(1,2,1,3,1,1,2,1))
refs %>% 
  mutate(numRawWords = str_count(text, "\\S+")) %>%
  group_by(objectSet) %>%
  tidyboot_mean(numRawWords) %>%
  left_join(scripted)
```

## TO-DO

1. Subject-level analyses of informativity: could be a mixture of 'lazy' egocentric speakers and 'attentive' audience-sensitive speakers? 
2. Sometimes the additional information is in movement instruction (i.e. "one square left toward a curtain" or "one square up below the comb")
3. Change in informativity over time (are people already being informative on first round? are people 'learning' to be more informative over course of expt? or getting less informative because tired of task?)
4. Just length of referring expression in words?

# Experiment 2

## Import turk survey data

```{r}
exp2_subjInfo <- read_csv('../data/final/experiment2/planned_sample/turk/subjInfo.csv')
```

## Import message data

```{r}
d_annotated <- read_csv('../data/final/experiment2/planned_sample/chatMessage/messages_with_annotations.csv')

# nonNativeSpeakerIDs <- unique((tangramSubjInfo %>% filter(nativeEnglish != "yes"))$gameid)
incompleteIDs <- unique((d_annotated %>% group_by(gameid) %>% 
                           filter(length(unique(trialNum)) != 24))$gameid)
confused <- unique((exp2_subjInfo %>% filter(understandsInstructions != 'yes'))$gameid)

# Some speakers relied on location in grid (TODO: WHAT ABOUT 'BELOW' PEOPLE?)
location_abusers <- c('8219-bb7861c1-43f4-480c-906b-8441c583c0ab', 
                      '7726-3d4a266f-e56d-4afb-9a79-b40deb690a76', 
                      '5684-f09de856-4d64-423c-b342-b30e7325eb0e', 
                      '7600-346d56b5-bd76-406f-befb-6b47e07d5916')
# Some speakers played this weird taboo game where they gave riddles 
tabooers <- c('2929-d218f724-b45e-416b-af44-5cab5658bad9', #(opposite of white... primary color...)
              '7600-346d56b5-bd76-406f-befb-6b47e07d5916', # MERRY GO ROUND
              '3462-7e95c955-4087-4eb9-bbc8-05338f9a5e4b') # CAN YOU SEE???
badGames <- c(incompleteIDs, location_abusers, confused)

d <- d_annotated %>%
  mutate(numFeatures = colorMention + shapeMention + textureMention) %>%
  filter(!(gameid %in% badGames))
```

## Check basic (no-hidden) pragmatic effect

```{r}
basic.d <- d %>%
  filter(hidden == 'no') %>%
  group_by(gameid, context) %>%
  summarize(m = mean(numRawWords)) %>%
  spread(context, m) %>%
  mutate(diff = close- far,
         ratio = log(close / far))

ggplot(basic.d, aes(x = far, y = close)) +
    geom_point() +
    theme_bw() +
    geom_abline(intercept = 0, slope = 1) +
    ylim(1,12) +
    xlim(1,12) +
    theme(aspect.ratio = 1, text = element_text(size=20))
```

Aggregated to difference scores:

```{r}
tidyboot_mean(basic.d %>%ungroup(), column = diff) %>%
  ggplot(aes(x = 'close - far', y = empirical_stat)) +
      geom_bar(stat ='identity', width = 0.25) +
      geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0) +
      theme_few()+
      geom_hline(yintercept = 0) +
      #geom_vline(xintercept = 0) +
      ylim(-1,1) +
      # xlim(-4,4) +
      ylab('# words used') +
    theme(aspect.ratio = 1, text = element_text(size=20))

t.test(basic.d$diff)
```

## Message length collapsing across all subjects

```{r}
dodge <- position_dodge(width=0.9)

d %>%
  group_by(context, hidden) %>%
  tidyboot_mean(column = numRawWords) %>%
  ggplot(aes(x = context, y = empirical_stat, fill = hidden)) +
    geom_bar(stat = 'identity', position = dodge) +
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), position = dodge, width = 0) +
    theme_few() +
    ylab('mean # words') +
    theme(aspect.ratio = 1, text = element_text(size=20)) +
    scale_fill_colorblind()

ggsave('~/Downloads/aggregated_means.pdf', width = 4, height = 2)
```

## Scatter to look at within-subj effects

Some people used exactly the same number of words on every trial...

```{r}
hidden.d %>% 
  filter(context == 'far') %>% 
  mutate(group = case_when(diff < -0 ~ 'anti',
                           diff > 0 ~ 'pro',
                           TRUE ~ 'all_the_things')) %>%
  group_by(group) %>% tally()

all_the_things_ers <- (hidden.d %>% filter(context =='far') %>% filter(diff == 0))$gameid
```

```{r}
hidden.d <- d %>%
  group_by(gameid, context, hidden) %>%
  summarize(m = mean(numRawWords)) %>%
  spread(hidden, m) %>%
  mutate(diff = yes- no,
         ratio = log(yes/no))

ggplot(hidden.d, aes(x = no, y = yes)) +
    geom_point() +
    facet_wrap(~ context) +
    theme_bw() +
    geom_abline(intercept = 0, slope = 1) +
    ylim(1,15) +
    ylab('# words (with occlusions)') +
    xlab('# words (without occlusions)') +
    #ggtitle('Within-pair comparisons of reference w/ and w/out occlusion') +
    xlim(1,15) +
    theme(aspect.ratio = 1, text = element_text(size=20))

ggsave('~/Downloads/scatter.pdf')
```

Histogram

```{r}
hidden.d %>% 
  #filter(!(gameid %in% all_the_things_ers)) %>% 
  group_by(context) %>% 
  mutate(m = mean(diff)) %>%
  ggplot(aes(x = diff)) +
      geom_histogram(binwidth = 0.33) +
      theme_bw() +
      geom_vline(xintercept = 0, linetype = 2) +
      geom_vline(aes(xintercept = m), color = 'red') +
      facet_wrap(~ context) +
      # ylim(1,12) +
      xlim(-6,6) +
      theme(aspect.ratio = 1)

ggsave("~/Downloads/histograms.pdf")
```

```{r}
hidden.d %>%
  group_by(context) %>%
  summarize(m = mean(diff), se = sd(diff)/sqrt(length(diff))) %>%
  ggplot(aes(x = context, y = m)) +
    geom_bar(stat = 'identity', width = 0.25) +
    geom_errorbar(aes(ymin = m - se, ymax = m + se), width = 0) +
    theme_few() +
    geom_hline(yintercept = 0) +
    ylim(-3, 3) +
    theme(aspect.ratio = 1, text = element_text(size=20))  +
    ylab('# words (occlusions - no occlusions)')

```

# Stats

Simple effect of hidden-ness in 'far' contexts

```{r}
summary(lmer(numRawWords ~ hidden + (1 + hidden | gameid), 
             data = d %>% 
               filter(context == 'far') %>%
               filter(!(gameid %in% all_the_things_ers))
               ))
t.test(numRawWords ~ hidden, data = d %>% filter(context == 'far') , var.equal = FALSE)
```

interaction-y way

```{r}
summary(lmer(numRawWords ~ context * hidden + (1 + context * hidden | gameid), data = d %>% filter(!(gameid %in% all_the_things_ers))))
```

helmert coding way (one way of slicing)

```{r}
d.regression <- d %>% 
  unite(condition, context, hidden) %>%
  ungroup() %>%
  mutate(condition = ordered(condition, levels = c('far_no', 'close_no', 'far_yes', 'close_yes')))
contrasts(d.regression$condition) <- contr.helmert(4)
summary(lmer(numRawWords ~ condition + (1 + condition | gameid), data = d.regression))
```

## Use feature data instead

```{r}
d %>% 
  group_by(context, occlusions) %>% 
  summarize(color = mean(colorMention), shape = mean(shapeMention), texture = mean(textureMention))
```

Histograms of number of features mentioned in each condition

```{r}
ggplot(d, aes(x = numFeatures)) +
  geom_histogram(binwidth = 1) +
  facet_grid(context ~ occlusions) +
  theme_few()
```

```{r}
d %>% group_by(context, occlusions) %>% 
  mutate(numTrials = length(text)) %>% 
  group_by(context, occlusions, colorMention, textureMention) %>% 
  summarize(n = length(text)/mean(numTrials)) %>%
  ggplot(aes(x = colorMention, y = textureMention, fill = n)) +
    geom_bin2d(stat = 'identity') +
    stat_bin2d(geom = "text", aes(label = round(n, 2))) +
    scale_fill_gradient(low = "white", high = "red") +
    facet_grid(occlusions ~ context) +
    theme_few()
```

Note: Random intercept doesn't converge (less variance in these numbers...)

```{r}
summary(lmer(numFeatures ~ context * occlusions + (1 + context + occlusions | gameid), data = d))
```

```{r}
summary(glmer(colorMention ~ occlusions + (1 + occlusions | gameid), family = 'binomial', data = d))
```

```{r}
summary(glmer(textureMention ~ occlusions + (1 + occlusions | gameid), family = 'binomial', data = d))
```

```{r}
d %>% 
  group_by(context, occlusions) %>% 
  summarize(numFeatures = mean(numFeatures))
```

```{r}
d.booted <- d %>% 
  gather(feature, mentioned, shapeMention, colorMention, textureMention) %>%
  group_by(context, hidden, feature) %>% 
  tidyboot_mean(column = mentioned)

ggplot(d.booted, aes(x = context, y = empirical_stat, fill = hidden)) + 
  geom_bar(stat = 'identity', position = dodge) +
  geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0) +
  facet_wrap(~ feature) + 
  theme_few() +
  theme(aspect.ratio = 1, text = element_text(size=20))  +
  ylab('% utterances with feature') +
  scale_fill_colorblind()

ggsave('~/Downloads/by-feature-analysis.pdf')
```