### 2018-06-27 analysis/final_analysis.Rmd
Add more model analyses, checking

---
title: "R Notebook"
output: html_notebook
---

# General imports

```{r}
library(jsonlite)
library(tidyverse)
library(ggthemes)
library(lme4)
library(lmerTest)
library(tidyboot)
```

# Experiment 1

```{r}
raw <- read_csv('../data/final/experiment1/chatMessage/messagesFromMongo.csv')
exp1_messages <- read_csv('../data/final/experiment1/chatMessage/messagesFromMongo.csv') %>%
  mutate(critical = ifelse(critical == 'True', 1, 0)) %>%
  rename(messageTime = serverTime) %>%
  filter(sender == 'director') %>%
  group_by(gameid, trialNum, instructionNum, attemptNum, critical, condition, objectSet, targetObject) %>%
  summarize(text = paste0(contents, collapse = ' '),
            typingRT = sum(typingRT)) %>%   # Concatenate multiple utterance on same trial
  mutate(numRawWords = str_count(text, "\\S+"))

## Have to post-hoc attemptNum and instructionNum bugs (got reset for next trial prematurely)
## Also, remove trials where multi-drop bug happened (i.e. where clicking multiple times triggered multiple instruction resetting)
exp1_drops <- read_csv('../data/final/experiment1/drop/dropsFromMongo.csv') %>%
  mutate(critical = ifelse(critical == 'True', 1, 0),
         correct = ifelse(correct == 'correct', 1, 0),
         instructionNum = ifelse(correct == 0, instructionNum + 1, instructionNum)) %>%
  group_by(gameid, targetObject) %>%
  filter(length(unique(instructionNum)) < 2) %>%
  group_by(gameid, instructionNum, trialNum) %>%
  mutate(attemptNum = ifelse(attemptNum == 0, max(attemptNum), attemptNum - 1)) %>%
  rename(dropTime = serverTime) %>%
  select(-X1, -eventType)
```

```{r}
exp1_subjInfo <- read_csv('../data/final/experiment1/turk/subjInfo.csv')
incompleteIDs <- unique((exp1_messages %>% 
                           group_by(gameid, trialNum, instructionNum) %>% 
                           summarize(id = row_number()) %>% 
                           group_by(gameid) %>% 
                           tally() %>% 
                           filter(n != 32))$gameid)

confused <- unique((exp1_subjInfo %>% filter(understandsInstructions != 'yes'))$gameid)
nonNative <- unique((exp1_subjInfo %>% filter(nativeEnglish != 'yes'))$gameid)

nonCriticalMistakes <- unique((exp1_drops %>%
  # only look at mistakes on noncritical (filler) items
  filter(critical != 1) %>%
  # don't want to double-count people for messing the same thing up multiple times
  filter(attemptNum == 1) %>% 
  group_by(gameid, correct) %>%
  tally() %>%
  # implement exclusion criteria of errors on >~10% of non-critical trials
  filter(n > 1))$gameid)

buggyTrials <- read_csv('../data/final/experiment1/drop/dropsFromMongo.csv') %>%
  mutate(critical = ifelse(critical == 'True', 1, 0),
       correct = ifelse(correct == 'correct', 1, 0),
       instructionNum = ifelse(correct == 0, instructionNum + 1, instructionNum)) %>%
  group_by(gameid, targetObject) %>%
  mutate(numInstructions = length(unique(instructionNum))) %>%
  filter(numInstructions > 1)

taboo <- c('4241-cda4a373-ffd7-41e3-ae60-82b9d377a455')
onlyLocation <- c('3222-f68d2df5-8977-4ff2-9ee9-c60f8ddb61f1', '9457-4189cc9b-a82b-4db9-828d-bf90d1987aee')
exp1_badGames <- c(incompleteIDs, nonNative, confused, nonCriticalMistakes, taboo, onlyLocation, unique(buggyTrials$gameid))

exp1_d <- exp1_drops %>% 
  left_join(exp1_messages) %>% #, by = c('gameid', 'trialNum', 'instructionNum', 'trialType', 
                      #         'targetObject', 'critical', 'condition', 'attemptNum')) %>%
  filter(!(gameid %in% exp1_badGames))
```

## Descriptive stats

How many recruited, how many excluded, for what reasons, etc?

```{r}
paste0('recruited ', length(exp1_subjInfo$gameid))
paste0('total games ', length(unique(exp1_drops$gameid)))
paste0('incomplete ', length(incompleteIDs))
paste0('too many mistakes ', length(setdiff(union(confused, nonCriticalMistakes), incompleteIDs)))
paste0('non-native english ', length(setdiff(nonNative, union(confused, union(nonCriticalMistakes, incompleteIDs)))))
paste0('final sample ', length(unique(exp1_d$gameid)))
```

Happened to get slightly more people in scripted condition

```{r}
exp1_d %>% 
  group_by(gameid, condition) %>% 
  tally() %>% 
  group_by(condition) %>% 
  summarize(numInCondition = length(n))
```

And slightly uneven numbers of 'experimental' trials in each item...

```{r}
exp1_d %>% 
  group_by(gameid, objectSet, trialType, condition) %>% 
  tally() %>% 
  group_by(condition, objectSet, trialType) %>% 
  summarize(numInCondition = length(n))
```

## Listener Analyses

First, look at error rates

```{r}
critTrials <- exp1_d %>% 
  filter(attemptNum == 0) %>%
  filter(critical == 1) %>%
  filter(trialType == 'exp') %>%
  mutate(error = 1 - correct)

critTrials %>%
  group_by(gameid, condition) %>%
  summarize(pctCriticalError = 4 - sum(correct)) %>%
  ggplot(aes(x = pctCriticalError, fill = condition)) +
    geom_histogram(alpha = 0.5, position = 'dodge', binwidth = 0.5)  +
    facet_grid(condition ~ .) +
    theme_few() +
    ylab('# pairs making k critical errors') +
    xlab('# critical errors') + 
    theme(aspect.ratio = 1, legend.position="none") +
    scale_fill_colorblind()+
    ggtitle('critical errors')

ggsave('../writing/journal_manuscript/figures/criticalErrors.pdf')
```

```{r}
critTrials %>%
  group_by(condition) %>%
  tidyboot_mean(column = error) %>%
  ggplot(aes(x = condition, y =empirical_stat, fill = condition)) +
    geom_bar(stat = 'identity', alpha = .5) +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0) +
    ylab('% errors') +
    scale_fill_colorblind()+
    theme_few() +
    theme(aspect.ratio = 1, legend.position="none")

ggsave('../writing/journal_manuscript/figures/errorMeans.pdf', height = 3, width = 3)
```

```{r}
errorTable <- critTrials %>%
  group_by(gameid, condition) %>%
  summarize(numErrors = 4 - sum(correct),
            numPossible = length(correct),
         atLeastOnce = sum(correct) < 4,
         atLeastTwice = sum(correct) < 3,
          atLeastThree = sum(correct) < 2,
         pctErrors = 1 - mean(correct)) %>%
  group_by(condition) %>%
  summarize(pctAtLeastOnce = mean(atLeastOnce),
            pctAtLeastTwice = mean(atLeastTwice),
            pctAtLeastThree = mean(atLeastThree),
            avgPctErrors = mean(pctErrors),
            totalNumErrors = sum(numErrors),
            totalPossible = sum(numPossible))
errorTable
```

### Statistical tests of error reduction

```{r}
prop.test(errorTable$totalNumErrors, errorTable$totalPossible)
```

### Control for item-level and subject-level error rates...

```{r}
summary(glmer(correct ~ condition + (1|gameid) + (1+condition | objectSet), family = 'binomial', 
              data = critTrials))
```

## Error heterogeneity across items...

```{r}
error_diffs <- critTrials %>% 
  group_by(objectSet, condition, targetObject) %>%
  summarize(pctErrors = mean(error)) %>%
  spread(condition, pctErrors) %>%
  mutate(errorReduction = scripted - unscripted) %>%
  gather(condition, pctErrors, scripted, unscripted) %>%
  right_join(critTrials) %>%
  group_by(objectSet, condition, targetObject, errorReduction) %>%
  tidyboot_mean(column = error)

ggplot(error_diffs, aes(x = reorder(targetObject, -errorReduction), y = empirical_stat, color = condition, group = condition)) +
    geom_point() +
    geom_line(stat = 'identity', position = 'dodge') +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0) +
    theme_few() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ggtitle('error heterogeneity across item') +
    xlab('critical item') +
    ylab('% error in condition')

ggsave('../writing/journal_manuscript/figures//itemHeterogeneity.pdf')
```

## Test error homogeneity

```{r}
errorRates = critTrials %>% 
  mutate(error = 1 - correct) %>%
  group_by(objectSet, condition, targetObject) %>%
  summarize(errors = sum(error), nonerrors = length(error) - errors) %>%
  ungroup()

scriptedErrors <- errorRates %>% 
  filter(condition == 'scripted') %>%
  select(errors, nonerrors)
chisq.test(as.matrix(rbind(scriptedErrors$errors, scriptedErrors$nonerrors)), simulate.p.value = T)

unscriptedErrors <- errorRates %>% 
  filter(condition == 'unscripted') %>%
  select(errors, nonerrors)
chisq.test(as.matrix(rbind(unscriptedErrors$errors, unscriptedErrors$nonerrors)), simulate.p.value = T)
```

```{r}
exp1_d %>% 
  filter(attemptNum == 0) %>%
  filter(correct == 1) %>%
  filter(critical == 1) %>% 
  ungroup() %>%
  filter(responseTime < median(responseTime, na.rm = T) +  3* sd(responseTime)) %>%
  group_by(trialType, condition) %>%
  mutate(m = median(responseTime)) %>%
  ggplot(aes(x = responseTime)) +
    geom_histogram() +
    geom_vline(aes(xintercept = m))+
    facet_wrap(trialType ~ condition) +
    scale_x_log10() +
    theme_few() 
```

```{r}
summary(lmer(log(responseTime) ~ trialType * condition + (1 | gameid),
     data = exp1_d %>% 
        filter(attemptNum == 0) %>%
        filter(correct == 1) %>%
        filter(critical == 1) %>% 
        ungroup() %>%
        filter(responseTime < median(responseTime, na.rm = T) +  3* sd(responseTime))))
```

## Mouse-tracking

Even on critical trials where they didn't make errors, how often did they 'consider' the distractor (i.e. hover over the distractor cell)

```{r}
d.onDistractor <- exp1_d %>% 
  left_join(read_csv('../data/final/experiment1/mouseOverDistractor/distractorMouseFromMongo.csv') %>% 
              mutate(critical = ifelse(critical == 'True', 1, 0))
            ) %>%
  select(-serverTime, -X1) %>%
  filter(attemptNum == 0) %>%
  filter(correct == 1) %>%
  filter(critical == 1) %>%
  group_by(gameid, trialNum) %>%
  mutate(toggle = floor((row_number() - 1) / 2)) %>%
  mutate(onOff = ifelse(is.na(onOff), 'off', onOff)) %>%
  spread(onOff, timeElapsed) %>%
  mutate(diff = ifelse(is.na(off), 0, off - on)) %>%
  group_by(gameid,condition,objectSet, trialType) %>%
  summarize(totalAmtTimeOnDistractor = sum(diff)) %>%
  mutate(hovered = totalAmtTimeOnDistractor > 0) %>%
  ungroup()

dodge <- position_dodge(width=0.9)
d.onDistractor %>%
  group_by(condition, trialType) %>%
  tidyboot_mean(column = hovered) %>%
  mutate(trial = ifelse(trialType == 'base', 'baseline', 'experimental')) %>%
  ggplot(aes(x = condition, y=empirical_stat, fill = condition, group = trial)) +
    geom_bar(aes(alpha = trial), stat = 'identity', position = dodge) +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0, position = dodge) +
    theme_few() +
    scale_alpha_manual(values=c(.25, .75)) +
    ylab('% trials hovering over distractor') +
    scale_fill_colorblind()+
    guides(fill='none') +
    theme(aspect.ratio = 1, legend.position = 'top')

ggsave('../writing/journal_manuscript/figures/hoverOverDistractor.pdf', height = 3, width = 3)
```

```{r}
summary(glmer(hovered ~ trialType * condition + (1 + trialType |gameid) + (1 |objectSet), data = d.onDistractor, family = 'binomial'))
```
Note that this is still significant for our preregistered dv (raw amount of data) -- it's just less interpretable given the distribution...

```{r}
summary(lmer(totalAmtTimeOnDistractor ~ trialType * condition + (1 + trialType|gameid) + (1 |objectSet), 
             data = d.onDistractor %>% mutate(totalAmtTimeOnDistractor = ifelse(totalAmtTimeOnDistractor == 0, 0, log(totalAmtTimeOnDistractor)))))
```

Note: because I stopped recording from the mouse as soon as someone starts dragging, it's technically possible for them to hover over the distractor object, click and start to drag it, but then drop it in place and go over to the target object. So they get it 'correct' even though it looks like their mouse never gets close...

Note: a couple researcher degrees of freedom here: 
* do we take the time from 'object reveal' or from first mouse move (i.e. subtract off the first time recorded NOT in target or distractor)? 
* do we restrict to only 'correct' trials (i.e. to make an argument like "even when they got it right, they were still distracted") or include everything?
* how big a radius around the target center should we consider a 'mouse-fixation' (obviously needs to be close enough to move it, but objects were different sizes...)

```{r}
# Touch people?
touch <- c('0095-03bb67a7-ae89-402d-9f7f-c5898784e63f', '0093-b4e1ce5c-2e5c-462a-bfed-87fa81611139')
targetRadius = 175
d.mouse <- exp1_d %>% ungroup() %>%
  left_join(read_csv('../data/final/experiment1/updateMouse/mouseFromMongo.csv') %>%
                  mutate(critical = ifelse(critical == 'True', 1, 0))
  ) %>%
  filter(!(gameid %in% touch)) %>%
  filter(critical == 1) %>% filter(attemptNum == 0) %>% filter(correct == 1) %>%
  select(-critical, -attemptNum, -eventType, -iterationName) %>%
  mutate(distractorDistance = as.numeric(distractorDistance)) %>%
  group_by(condition, trialType, gameid, objectSet, onTarget = targetDistance < targetRadius) %>%
  summarize(firstTime = first(timeFromReveal), 
            lastTime = last(timeFromReveal)) %>%
  group_by(condition, trialType, gameid, objectSet) %>%
  summarize(initMouseMove = min(firstTime), 
            initTargetHover = max(firstTime) - initMouseMove, 
            finalTargetHover = max(lastTime) - initMouseMove) %>%
  select(-initMouseMove) %>%
  gather(metric, value, initTargetHover, finalTargetHover) %>%
  #spread(onTarget, value) %>%
  #mutate(timeFromFirstMouseMove = `TRUE` - `FALSE`) %>%
  #ungroup() %>%
  #filter(firstTime < mean(firstTime, na.rm = T) + 3 * sd(firstTime, na.rm = T)) %>%
  group_by(condition, trialType, metric) %>%
  tidyboot_mean(column = value)

d.mouse
ggplot(d.mouse, aes(x = trialType, y = empirical_stat, fill = metric)) +
  geom_bar(stat= 'identity', position = dodge) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), position = dodge, width = 0) +
  facet_wrap(~ condition) +
  theme_few()
```

## Prep informativity norming stims

```{r}
refExps <- exp1_d %>% 
  filter(condition == 'unscripted') %>% 
  filter(critical == 1) %>% 
  filter(attemptNum == 0) %>% 
  ungroup() %>%
  select(gameid, objectSet, text)

write_csv(refExps, './refExpressions_raw.csv')
```

Look at all of them to make sure there aren't dumb duplicates...

```{r}
scripted_words <- c('glasses', 'bottom block', 'tape', 'large measuring cup', 'brush', 'eraser', 'small candle', 'mouse')
scripted_refs <- data.frame(objectSet = c(1,2,3,4,5,6,7,8), text = scripted_words, stringsAsFactors=FALSE)

standardized_refs <- read_csv('../data/final/experiment1/refExpressions_standardized.csv')
cat('unique utterances: ', length(unique(c(scripted_refs, standardized_refs$text))))
unscripted_utt_counts <- standardized_refs %>% 
  group_by(objectSet,text) %>%
  tally() %>%
  mutate(condition = 'unscripted')
```

Unscripted speakers had significantly more words in their referring expressions across items... (note: if you just bootstrap cis for each item, it's significant for 6 of 8...)

```{r}
unscripted_refs <- read_csv('../data/final/experiment1/refExpressions.csv')
scripted.df <- data.frame(objectSet = c(1,2,3,4,5,6,7,8), scriptedMessage = scripted_refs, scriptedNumWords = c(1,2,1,3,1,1,2,1))
diff.df <- unscripted_refs %>% 
  mutate(unscriptedNumWords = str_count(text, "\\S+")) %>%
  left_join(scripted.df, by = c('objectSet')) %>%
  mutate(diff = unscriptedNumWords - scriptedNumWords)

library(lmerTest)
summary(lmer(diff ~ 1 + (1 | objectSet) + (1 | gameid), data= diff.df))
```

## Analyze informativity ratings

First identify 4 turkers who were responding randomly...

```{r}
raw_inf_ratings <- read_csv('../informativity_ratings.csv')
# ggplot(raw_inf_ratings %>% mutate(workerid = substr(workerid, 0, 6)), aes(x = response, fill = referent)) +
#   geom_density(alpha = 0.5) +
#   facet_wrap(~ workerid, scales = 'free_y')
# 
# ggplot(raw_inf_ratings %>% mutate(workerid = substr(workerid, 0, 6)), aes(x = slideNumber, y = response, color = referent)) + 
#   geom_point() +
#   facet_wrap(~ workerid)

raw_inf_ratings %>% mutate(workerid = substr(workerid, 0, 6)) %>%
  group_by(workerid) %>%
  summarize(medianRT = median(rt))

bad_raters <- c('956e201e2604ead2f2fa81d0ef666626', '6559a18fae791987122350d706538e90', '2301c2d66270fecb3c16cd1593e4aa0f', 'dd0463e01b62413231a9218253908a91')#, 'deec4028cd9f327085a947bac0e19aab', '766e6aa74bc2c113e29531dab30d63e9')
```

```{r}
library(irr)
inf_ratings <- raw_inf_ratings %>%
  filter(!(workerid %in% bad_raters))

inf_diffs <- inf_ratings %>%
  select(-slideNumber, -rt) %>%
  group_by(workerid, label) %>%
  spread(referent, response) %>%
  mutate(diff = target - distractor) %>%
  ungroup()

wideRatings <- inf_diffs %>% select(label, diff, workerid) %>% spread(workerid, diff)
icc(wideRatings[,-c(1)],
    model = "twoway", type = "agreement")

inf_diffs_collapsed <- inf_diffs %>%
  group_by(label) %>%
  summarize(distractor = mean(distractor),
            target = mean(target),
            paired_inf = mean(diff))
```

lmer mixed-model here is complicated because of the complex & partially crossed nature of the variance in the data (i.e. between judgements of raters, between object sets, between utterances chosen by different speakers)  TODO: double-bootstrapping thing where you first resample 'mean' ratings for each utterance, then you resample the distribution of utterances in a condition (i.e. within object set), then you take a weighted average to get your 'informativity'

```{r}
## Make a df with each utterance duplicated proportionally to the times it was said
utt_data_tmp <- scripted_refs %>% mutate(n = 68) %>% mutate(condition = 'scripted') %>%
  group_by(text) %>% 
  slice(rep(1:n(), each = n)) %>%
  mutate(gameid = as.character(row_number()))

utt_data <- left_join(
  inf_diffs %>% group_by(label) %>% summarize(m = mean(diff)),
  utt_data_tmp %>% rename(label = text)
) %>% filter(condition == 'scripted') %>% group_by(label, objectSet, condition) %>% summarize(m = mean(m)) %>%
  right_join(critTrials %>% filter(condition == 'scripted'), by = c('objectSet', 'condition')) %>%
  group_by(gameid, condition, objectSet, label) %>%
  group_by(label, objectSet, condition, gameid) %>%
  tally() %>%
  select(-n) %>%
  rename(text = label) %>%
  bind_rows(standardized_refs %>% mutate(condition = 'unscripted', objectSet = as.numeric(objectSet)))
```

```{r}
## In the inner loop, we want to sample a 'mean' rating for each utterance
## using same sample of *raters* for all utterances... 
sampleRatings = function(df, indices) {
  grouped_df <- df %>%
    group_by(label, objectSet) %>%
    do(data.frame(diff = mean(.$diff[indices]),
                  target = mean(.$target[indices]),
                  distractor = mean(.$distractor[indices])))
  return(grouped_df)
}

## In outer loop, we want to sample 'pairs' as a unit (with their 8 utterances)
samplePairs = function(df) {
  uniq_ids <- unique(df$gameid)
  new_ids <- sample(uniq_ids, length(uniq_ids), replace = TRUE)
  resampled_ids = data.frame(gameid = new_ids,
                             sampleid = 1:length(new_ids), 
                             stringsAsFactors = FALSE)
  return(left_join(resampled_ids, df, by = c('gameid')))
}

## Finally, want to sample object sets per pair
sampleObjects = function(df, indices) {
  new_objects <- data.frame(objectSet = df$objectSet[indices])
  return(left_join(new_objects, df, by = c('objectSet')))
}
# sampleObjects(data.frame(objectSet = c(1,2,3,4,5,6,7,8), utt = c('a', 'b', 'c', 'a', 'b', 'c', 'd', 'e')),
#               obj_sets)
```

```{r}
## do a multi-stage bootstrap...
## 1. resample a set of raters to get a mean for each utterance...
## 2. sample object set indices (e.g. 5,2,3,3,5,1,8,5)
## 3. resample pairs within each condition and then resample utterances within pairs using object set indices
boot_samples = replicate(1000, {
  raters = sample(1:16,16,replace = TRUE)
  obj_sets = sample(1:8, 8, replace = TRUE)
  resampled_ratings = sampleRatings(inf_diffs, raters)

  resampled_pairs = utt_data %>%
    group_by(condition) %>%
    do(samplePairs(.)) %>%
    group_by(condition, sampleid) %>%
    do(sampleObjects(., obj_sets)) %>%
    rename(label = text)

  combined <- left_join(resampled_pairs, resampled_ratings, by = 'label') %>%
    gather(referent, value, target,distractor,diff) %>%
    group_by(condition, referent) %>%
    summarize(m = mean(value))
  diff_of_diffs <- combined %>% filter(referent == 'diff') %>% spread(condition, m) %>% mutate(diff_of_diffs = unscripted - scripted)
  return(c((combined %>% filter(referent != 'diff'))$m, diff_of_diffs$diff_of_diffs))
})
```

```{r}
# Look at diff of diffs (if excludes 0, significant)
data.frame(test_stat = boot_samples[5,]) %>%
  summarize(m = mean(test_stat),
            ci_upper = sort(test_stat)[length(test_stat) * 0.025],
            ci_lower = sort(test_stat)[length(test_stat) * 0.975]) 
```

```{r}
boot_stats = data.frame(
  scripted_distractor = boot_samples[1,],
  scripted_target = boot_samples[2,],
  unscripted_distractor = boot_samples[3,],
  unscripted_target = boot_samples[4,]
) %>%
  gather(condition, value) %>%
  group_by(condition) %>%
  summarize(m = mean(value),
            ci_upper = sort(value)[length(value) * 0.025],
            ci_lower = sort(value)[length(value) * 0.975]) %>%
  separate(condition, into = c('condition', 'referent'), by = '_')

dodge <- position_dodge(width=0.9)
ggplot(boot_stats, aes(x = condition, y = m, fill = condition, alpha = referent)) +
  geom_bar(stat= 'identity', position = 'dodge') +
  scale_alpha_manual(values=c(.25, .75)) +
  geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0, position = dodge) + 
  theme_few() +
  scale_fill_colorblind() +
  guides(fill = 'none') +
  ylab('utterance informativity ratings') +
  theme(aspect.ratio = 1)

ggsave('../writing/journal_manuscript/figures/informativity.pdf')
```

### How well does informativity correlate with error?

now bootstrap again but within items

```{r}
boot_objectsets = replicate(1000, {
  raters = sample(1:16,16,replace = TRUE)
  resampled_ratings = sampleRatings(inf_diffs, raters)

  resampled_pairs = utt_data %>%
    group_by(condition) %>%
    do(samplePairs(.)) %>%
    rename(label = text)

  combined <- left_join(resampled_pairs, resampled_ratings, by = c('label', 'objectSet')) %>%
    gather(referent, value, target,distractor,diff) %>%
    group_by(condition, objectSet, referent) %>%
    summarize(m = mean(value))

  return((combined %>% filter(referent == 'diff'))$m)
})
```

Kind of messy way to aggregate these, but so it goes... (replicate doesn't return a dataframe)

(note that the 'outlier' here is the scripted 'candle' trial: raters thought it was *much* more likely to refer )

```{r}
boot_objectsets_stats = data.frame(
  condition = c(rep('scripted', 8), rep('unscripted', 8)),
  m_rating = rep(0,16),
  ci_upper_rating = rep(0,16),
  ci_lower_rating = rep(0,16),
  objectSet = rep(1:8,2)
)

for (i in 1:16) {
  boot_row = boot_objectsets[i,]
  boot_objectsets_stats[i,]$m_rating = mean(boot_row)
  boot_objectsets_stats[i,]$ci_upper_rating = sort(boot_row)[length(boot_row) * 0.025]
  boot_objectsets_stats[i,]$ci_lower_rating = sort(boot_row)[length(boot_row) * 0.975]
}

left_join(boot_objectsets_stats, error_diffs, by = c('condition', 'objectSet')) %>%
  ggplot(aes(x = m_rating, y = mean, color = condition, group = 1)) +
    geom_point(size = 2, shape = 16) +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), alpha = .3) +
    geom_errorbarh(aes(xmax = ci_upper_rating, xmin = ci_lower_rating), alpha =.3) +
    geom_smooth(linetype = 'dotted', method = 'lm', colour="black", se = F) +
    theme_few() +
    theme(aspect.ratio = 1, legend.position = 'top') +
    scale_color_colorblind() +
    xlab('informativity diff (target - distractor)') +
    ylab('% critical error')

ggsave('../writing/journal_manuscript/figures/informativity_predicts_errors.pdf',
       height = 3, width = 3, useDingbats=FALSE)
```

```{r}
boot_correlation = replicate(1000, {
  raters = sample(1:16,16,replace = TRUE)
  resampled_ratings = sampleRatings(inf_diffs, raters)

  resampled_pairs = utt_data %>%
    group_by(condition) %>%
    do(samplePairs(.)) %>%
    rename(label = text)

  combined <- left_join(resampled_pairs, resampled_ratings, by = c('label', 'objectSet')) %>%
    select(-target,-distractor,-label) %>%
    inner_join(critTrials, by = c('gameid', 'condition', 'objectSet')) %>%
    group_by(objectSet, condition) %>%
    summarize(inf = mean(diff), error = mean(error))

  return(cor(combined$inf, combined$error))
})

cat('r = ', mean(boot_correlation), 
    ', 95% CI = [', sort(boot_correlation)[length(boot_correlation) * 0.025],
    ', ', sort(boot_correlation)[length(boot_correlation) * 0.975],']')
```

## TO-DO

1. Subject-level analyses of informativity: could be a mixture of 'lazy' egocentric speakers and 'attentive' audience-sensitive speakers? 
2. Sometimes the additional information is in movement instruction (i.e. "one square left toward a curtain" or "one square up below the comb")
3. Change in informativity over time (are people already being informative on first round? are people 'learning' to be more informative over course of expt? or getting less informative because tired of task?)
4. Order effects (i.e. are people more likely to use a modifier when they had to on the previous round?)
5. Are people more informative when there are more objects?

# Experiment 2

## Import turk survey data

```{r}
exp2_subjInfo <- read_csv('../data/final/experiment2/planned_sample/turk/subjInfo.csv')
```

## Import message data

```{r}
d_annotated <- read_csv('../data/final/experiment2/planned_sample/chatMessage/messages_with_annotations.csv')

# nonNativeSpeakerIDs <- unique((tangramSubjInfo %>% filter(nativeEnglish != "yes"))$gameid)
incompleteIDs <- unique((d_annotated %>% group_by(gameid) %>% 
                           filter(length(unique(trialNum)) != 24))$gameid)
confused <- unique((exp2_subjInfo %>% filter(understandsInstructions != 'yes'))$gameid)
nonNative <- unique((exp2_subjInfo %>% filter(nativeEnglish != 'yes'))$gameid)

# Some speakers relied on location in grid (TODO: WHAT ABOUT 'BELOW' PEOPLE?)
location_abusers <- c('8219-bb7861c1-43f4-480c-906b-8441c583c0ab', 
                      '7726-3d4a266f-e56d-4afb-9a79-b40deb690a76', 
                      '5684-f09de856-4d64-423c-b342-b30e7325eb0e', 
                      '7600-346d56b5-bd76-406f-befb-6b47e07d5916')
# Some speakers played this weird taboo game where they gave riddles 
tabooers <- c('2929-d218f724-b45e-416b-af44-5cab5658bad9', #(opposite of white... primary color...)
              '7600-346d56b5-bd76-406f-befb-6b47e07d5916', # MERRY GO ROUND
              '3462-7e95c955-4087-4eb9-bbc8-05338f9a5e4b') # CAN YOU SEE???
badGames <- c(incompleteIDs, nonNative, location_abusers, confused, tabooers)

d <- d_annotated %>%
  mutate(numFeatures = colorMention + shapeMention + textureMention) %>%
  filter(!(gameid %in% badGames)) %>%
  mutate(occlusions = as.factor(occlusions),
        context = as.factor(context))
```

Write out for BDA

```{r}
d %>% 
  select(-numRawWords, -numFeatures, -text, -occlusions) %>%
  write_csv('../modeling/bdaInput/bdaInput.csv')
```

## Descriptive stats

How many recruited, how many excluded, for what reasons, etc?

```{r}
paste0('recruited ', length(exp2_subjInfo$gameid))
paste0('total games ', length(unique(d_annotated$gameid)))
paste0('incomplete ', length(incompleteIDs))
paste0('confused ', length(setdiff(confused,incompleteIDs)))
paste0('non-native english ', length(setdiff(nonNative, union(confused, incompleteIDs))))
paste0('violated instructions ', length(setdiff(c(location_abusers, tabooers), union(nonNative, union(confused, incompleteIDs)))))
paste0('final sample ', length(unique(d$gameid)))
```

## Quick listener analysis: how common are errors?

Errors rare...

```{r}
exp2_clicks <- read_tsv('../data/final/experiment2/planned_sample/clickedObj/clickedObj.csv') %>%
  mutate(correct = ifelse(correct == 'true', 1, 0)) %>%
  filter(!(gameid %in% badGames))

## Errors by condition
tmp <- exp2_clicks %>%
  group_by(context, occlusions) %>%
  summarize(m = sum(correct), total = length(correct) - m)

chisq.test(t(tmp[,3:4]))

## Error distribution by pair
exp2_clicks %>%
  group_by(gameid) %>%
  summarize(numErrors = 24- sum(correct)) %>%
  group_by(numErrors) %>%
  tally()
```

Add missing column annotating whether critical distractor in close trials different from target in texture, color, or both...

```{r}
library(dplyr)
bdaInput = exp2_clicks %>% 
  mutate(names = substr(names,2,nchar(names)-1)) %>%
  separate(names, sep = ',', into = c('item1','item2','item3', 'item4', 'item5'),
           extra = "merge", fill = "left") %>%
  select(-item5) %>%
  gather(distractorNum, distractorName, item1:item4) %>%
  separate(intendedName, into = c('targetTexture', 'targetColor', 'targetShape')) %>%
  filter(!is.na(distractorName)) %>%
  separate(distractorName, into = c('distractorTexture', 'distractorColor', 'distractorShape')) %>%
  filter(distractorShape == targetShape) %>%
  mutate(textureMatch = distractorTexture == targetTexture) %>%
  mutate(colorMatch = distractorColor == targetColor) %>%
  mutate(distractorType = ifelse(textureMatch, 'texture_shape', ifelse(colorMatch, 'color_shape', 'shape_only'))) %>%
  select(gameid, trialNum, distractorType) %>%
  right_join(exp2_clicks) %>%
  mutate(distractorType = ifelse(is.na(distractorType), 'diff_shape', distractorType)) %>%
  right_join(d) %>%
  select(gameid, trialNum, distractorType, context, hidden, numDistractors, colorMention, shapeMention, textureMention)

write_csv(bdaInput, '../modeling/bdaInput/bdaInput.csv')
```

Now that we have these distractor types... check if results differ across what the distractor is?

```{r}
bdaInput %>%
  group_by(distractorType, hidden) %>%
  tidyboot_mean(column = numFeatures) %>%
  ggplot(aes(x = distractorType, y = mean, fill = hidden)) +
    geom_bar(stat = 'identity', position = dodge) +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0)
```

check if results depend on number of distractors?

```{r}
bdaInput %>%
  mutate(numVisible= numDistractors - numObjsOccluded) %>%
  group_by(numVisible, hidden) %>%
  tidyboot_mean(column = numFeatures) %>%
  ggplot(aes(x = numVisible, y = mean, fill = hidden)) +
    geom_bar(stat = 'identity', position = dodge) +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0)
```

## Visualize basic (no-hidden) pragmatic effect

```{r}
basic.d <- d %>%
  filter(hidden == 'no') %>%
  group_by(gameid, context) %>%
  summarize(m = mean(numRawWords)) %>%
  spread(context, m) %>%
  mutate(diff = close- far,
         ratio = log(close / far))

summary(lmer(numRawWords ~ context + (1 + context | gameid), d %>% filter(hidden == 'no')))

ggplot(basic.d, aes(x = far, y = close)) +
    geom_point() +
    theme_bw() +
    geom_abline(intercept = 0, slope = 1) +
    ylim(1,12) +
    xlim(1,12) +
    theme(aspect.ratio = 1, text = element_text(size=20))
```

Aggregated to difference scores:

```{r}
tidyboot_mean(basic.d %>%ungroup(), column = diff) %>%
  ggplot(aes(x = 'close - far', y = empirical_stat)) +
      geom_bar(stat ='identity', width = 0.25) +
      geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0) +
      theme_few()+
      geom_hline(yintercept = 0) +
      #geom_vline(xintercept = 0) +
      ylim(-1,1) +
      # xlim(-4,4) +
      ylab('# words used') +
    theme(aspect.ratio = 1, text = element_text(size=20))

t.test(basic.d$diff)
```

## Visualize message length collapsing across all subjects

```{r}
dodge <- position_dodge(width=0.9)
cbPalette = c(rgb(0.8, .8, .8), rgb(.8, .3, .3),  rgb(0, .6, .4), rgb(.4, .4, .4))

d %>%
  group_by(context, hidden) %>%
  tidyboot_mean(column = numRawWords) %>%
  ggplot(aes(x = context, y = empirical_stat, fill = hidden)) +
    geom_bar(stat = 'identity', position = dodge) +
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), position = dodge, width = 0) +
    theme_few() +
    ylab('mean # words') +
    theme(aspect.ratio = 1, text = element_text(size=20)) +
    scale_fill_manual(values=cbPalette) +
    theme(aspect.ratio = 1)

ggsave('../writing/journal_manuscript/figures/num_words.pdf', width = 4, height = 3)
```

## Visualize individual differences

Some people used exactly the same number of words on every trial...

```{r}
hidden.d <- d %>%
  group_by(gameid, context, hidden) %>%
  summarize(m = mean(numRawWords)) %>%
  spread(hidden, m) %>%
  mutate(diff = yes- no,
         ratio = log(yes/no)) 
```

```{r}
ggplot(hidden.d, aes(x = no, y = yes)) +
    geom_point() +
    facet_wrap(~ context) +
    theme_bw() +
    geom_abline(intercept = 0, slope = 1) +
    ylim(1,15) +
    ylab('# words (with occlusions)') +
    xlab('# words (without occlusions)') +
    #ggtitle('Within-pair comparisons of reference w/ and w/out occlusion') +
    xlim(1,15) +
    theme(aspect.ratio = 1, text = element_text(size=20))

ggsave('../writing/journal_manuscript/figures//scatter.pdf')
```

Histogram

```{r}
hidden.d %>% 
  #filter(!(gameid %in% all_the_things_ers)) %>% 
  group_by(context) %>% 
  mutate(m = mean(diff)) %>%
  ggplot(aes(x = diff)) +
      geom_histogram(binwidth = 0.33) +
      theme_bw() +
      geom_vline(xintercept = 0, linetype = 2) +
      geom_vline(aes(xintercept = m), color = 'red') +
      facet_wrap(~ context) +
      # ylim(1,12) +
      xlim(-6,6) +
      theme(aspect.ratio = 1)

ggsave("~/Downloads/histograms.pdf")
```

Look for population who mentioned all the things all the time...

```{r}
hidden.d %>% 
  filter(context == 'far') %>% 
  mutate(group = case_when(diff < -0 ~ 'anti',
                           diff > 0 ~ 'pro',
                           TRUE ~ 'all_the_things')) %>%
  group_by(group) %>% tally()

all_the_things_ers <- (hidden.d %>% filter(context =='far') %>% filter(diff == 0))$gameid
```

```{r}
hidden.d %>%
  group_by(context) %>%
  summarize(m = mean(diff), se = sd(diff)/sqrt(length(diff))) %>%
  ggplot(aes(x = context, y = m)) +
    geom_bar(stat = 'identity', width = 0.25) +
    geom_errorbar(aes(ymin = m - se, ymax = m + se), width = 0) +
    theme_few() +
    geom_hline(yintercept = 0) +
    ylim(-3, 3) +
    theme(aspect.ratio = 1, text = element_text(size=20))  +
    ylab('# words (occlusions - no occlusions)')

```

# Stats

Simple effect of hidden-ness in 'far' contexts

```{r}
# summary(lmer(numRawWords ~ hidden + (1 + hidden | gameid), 
#              data = d %>% filter(context == 'far') 
))
#t.test(numRawWords ~ hidden, data = d %>% filter(context == 'far') , var.equal = FALSE)
```

interaction-y way

```{r}
contrasts(d$context) <- contr.treatment(2, base = 2)
contrasts(d$occlusions) <- contr.treatment(2, base = 2)

summary(lmer(numRawWords ~ context * occlusions + (1 + context * occlusions | gameid), data = d ))
```

## Use feature data instead

```{r}
d %>% 
  group_by(context, occlusions) %>% 
  summarize(color = mean(colorMention), shape = mean(shapeMention), texture = mean(textureMention))
```

Histograms of number of features mentioned in each condition

```{r}
ggplot(d, aes(x = numFeatures)) +
  geom_histogram(binwidth = 1) +
  facet_grid(context ~ occlusions) +
  theme_few()
```

```{r}
d %>% group_by(context, occlusions) %>% 
  mutate(numTrials = length(text)) %>% 
  group_by(context, occlusions, colorMention, textureMention) %>% 
  summarize(n = length(text)/mean(numTrials)) %>%
  ggplot(aes(x = colorMention, y = textureMention, fill = n)) +
    geom_bin2d(stat = 'identity') +
    stat_bin2d(geom = "text", aes(label = round(n, 2))) +
    scale_fill_gradient(low = "white", high = "red") +
    facet_grid(occlusions ~ context) +
    theme_few()
```

Note: Random interaction doesn't converge (less variance in these numbers...)

```{r}
```

```{r}
contrasts(d$context) <- contr.treatment(2, base = 2)
contrasts(d$occlusions) <- contr.treatment(2, base = 2)
summary(glmer(colorMention ~ occlusions * context + (1 + occlusions | gameid), 
              family = 'binomial', data = d))
```

```{r}
summary(glmer(textureMention ~ occlusions * context + (1 + occlusions | gameid), 
              family = 'binomial', data = d))
summary(tmp2)
fixef(tmp2)
baseprob = fixef(tmp2)[1] 
cat('no-occlusion:', exp(baseprob) / (1 + exp(baseprob)))
occludedprob = fixef(tmp2)[1] + fixef(tmp2)[2]
cat('with occlusion:', exp(occludedprob) / (1 + exp(occludedprob)))
```

```{r}
d %>% 
  group_by(context, hidden) %>% 
  tidyboot_mean(column = numFeatures) %>%
  ggplot(aes(x = context, y = mean, fill = hidden)) +
    geom_bar(stat = 'identity', position = dodge) +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0)
```

```{r}
d.booted <- d %>% 
  gather(feature, mentioned, shapeMention, colorMention, textureMention) %>%
  group_by(context, hidden, feature) %>% 
  tidyboot_mean(column = mentioned)

ggplot(d.booted, aes(x = context, y = mean, fill = hidden)) + 
  geom_bar(stat = 'identity', position = dodge) +
  geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0) +
  facet_wrap(~ feature) + 
  theme_few() +
  theme(aspect.ratio = 1, text = element_text(size=20))  +
  ylab('% utterances with feature') +
  scale_fill_manual(values=cbPalette)

ggsave('../writing/journal_manuscript/figures/by-feature-analysis.pdf', 
       height = 3, width = 9)
```

## Model-based analyses

```{r}
library(coda)

estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}
HPDhi<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}
HPDlo<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}

rbind(read.csv('../modeling/bdaOutput/occlusionSensitivityPredictives.csv') %>% mutate(source = 'occlusion-sensitive model'),
      read.csv('../modeling/bdaOutput/egocentricPredictives.csv') %>% mutate(source = 'egocentric model')) %>%
  mutate(i = row_number()) %>%
  gather(condition, value, close_hidden:far_visible) %>%
  group_by(condition, source) %>%
  summarize(mean = estimate_mode(value),
            ci_lower = round(HPDlo(value), 3),
            ci_upper = round(HPDhi(value), 3)) %>%
  separate(condition, into = c('context', 'hidden')) %>%
  mutate(hidden = ifelse(hidden == 'hidden', 'yes', 'no')) %>%
  rbind(d %>% 
          group_by(context, hidden) %>% 
          tidyboot_mean(column = numFeatures) %>%
          ungroup() %>%
          mutate(context = as.character(context)) %>%
          mutate(source = as.character('empirical data')) %>% 
          select(-empirical_stat, -n)) %>%
  ggplot(aes(x = context, fill = hidden, y = mean)) +
    geom_bar(stat = 'identity', position = dodge) +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0) +
    facet_wrap(~ source) +
    theme_few() +
    ylab('mean # features produced') +
    theme(aspect.ratio = 1)
```

Examine param posteriors for occlusion-sensitive model

```{r}
read.csv('../modeling/bdaOutput/occlusionSensitivityParams.csv') %>%
  gather(parameter, value, alpha:shapeCost) %>%
  group_by(parameter) %>%
  summarize(mode = estimate_mode(value),
            md_lo = round(HPDlo(value), 3),
            md_hi = round(HPDhi(value), 3))
```

log-likelihood plot

```{r}

```